Dynamic strategies for yaw and induction control of wind farms based on large-eddy simulation and optimization

### 2018-01-02 reply_to_reviewer1.tex

\paragraph{Response}

\hrulefill

\paragraph{4. Reviewer} \textit{Line 291-292. It would be beneficial to include the turbulent case also, making it a 4-panel figure. That case is more representative of an ABL flow and therefore provides the reader a better sense of the optimization convergence in a more realistic scenario.}

\paragraph{Response} As indicated in the text, the turbulent inflow case follows a similar behavior as the uniform inflow. Including both uniform and turbulent convergence characteristics would in our opinion draw the attention away from the more important results in the paper. As a compromise, we replaced the uniform optimization convergence with the turbulent one in the revised manuscript.

### 2018-01-02 MuntersMeyers_DynamicStrategiesControl_revision_WM01.tex
R1C4

\subsection{Convergence behavior}\label{sec:opt_convergence}
\noindent Figure~\ref{fig:convergence} illustrates the convergence behavior within the second optimization window of the uniform inflow case. Figure~\ref{fig:convergence}a shows that, although formal convergence is not achieved for any of the cases, after 80 iterations, the relative ordering of the cases is clear and we assume that further improvements in the cost function for an increased amount of iterations are marginal. From Figure~\ref{fig:convergence}b it can be seen that the amount of PDE evaluations (forward or adjoint) is approximately twice the amount of BFGS iterations, indicating that the optimizer mostly takes Newton steps without a further line search. The strong increase in PDE evaluations for I3 and I2Y can be explained by the fact that the limits of continuous adjoint gradient accuracy complicate satisfaction of the Wolfe conditions upon approaching a (local) optimum. Similar characteristics were observed for the turbulent case (not further shown here).

\begin{figure}
	\includegraphics[width=\textwidth]{figure4}
	\caption{Convergence behavior for the second optimization window in uniform inflow. \emph{a) }: Cost functional $\mathscr{J}$. \emph{b) }: PDE evaluations (forward or adjoint). Markers represent every fifteenth iteration step.\label{fig:convergence}}
\end{figure}

>>>>

\subsection{Convergence behavior}\label{sec:opt_convergence}
\noindent Figure~\ref{fig:convergence} illustrates the convergence behavior within the second optimization window of the \revision{turbulent} inflow case. Figure~\ref{fig:convergence}a shows that, although formal convergence is not achieved for any of the cases, after 80 iterations, the relative ordering of the cases is clear and we assume that further improvements in the cost function for an increased amount of iterations are marginal. From Figure~\ref{fig:convergence}b it can be seen that the amount of PDE evaluations (forward or adjoint) is approximately twice the amount of BFGS iterations, indicating that the optimizer mostly takes Newton steps without a further line search. Similar characteristics were observed for the \revision{uniform} case (not further shown here).

\begin{figure}
	\includegraphics[width=\textwidth]{figure4_revised}
	\caption{Convergence behavior for the second optimization window in \revision{turbulent} inflow. \emph{a) }: Cost functional $\mathscr{J}$. \emph{b) }: PDE evaluations (forward or adjoint). Markers represent every fifteenth iteration step.\label{fig:convergence}}
\end{figure}
